---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# **blockpop**: Estimate Census Block Populations for 2020

<!-- badges: start -->
[![R-CMD-check](https://github.com/CoryMcCartan/blockpop/workflows/R-CMD-check/badge.svg)](https://github.com/CoryMcCartan/blockpop/actions)
[![Lifecycle: experimental](https://img.shields.io/badge/lifecycle-experimental-orange.svg)](https://lifecycle.r-lib.org/articles/stages.html#experimental)
<!-- badges: end -->

2020 Census data is delayed and will be affected by differential privacy.
This package uses FCC block-level population estimates from 2010--2019, which
are based on new roads and map data, to estimate 2020 block populations.  In
the future it may estimate other quantities of interest as well.

## Installation

You can install the development version from [GitHub](https://github.com/) with:

``` r
# install.packages("devtools")
devtools::install_github("CoryMcCartan/blockpop")
```

## Usage

We start by downloading the FCC data locally (although you can skip this
step if you just want estimates for one state).

```{r include=F}
library(blockpop)
```

```{r eval=F}
library(blockpop)
download_fcc("data-raw/fcc.csv")
```

Then we can extract the data for the state we care about and construct the 2020 
block estimates.

```{r message=F}
library(dplyr)

WA_blocks = load_state("WA", "data-raw/fcc.csv")
WA_est = est_2020(WA_blocks)

print(WA_est)
summarize(WA_est, across(starts_with("pop"), sum))
```
