---
title: "ACS-Decennial-Block Harmonization"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{ACS-Decennial-Block Harmonization}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include=FALSE}
knitr::opts_chunk$set(collapse=TRUE, comment = "#>",
                      eval=file.exists(here::here("data-raw/fcc.csv")))
```

This vignette examines the accuracy of the ACS-Decennial-Block harmonization
calculated by `harmonize_vars()`. We start where the [README](../) leaves off.

```{r eval=F}
library(blockpop)
library(dplyr)
library(stringr)
library(tidyr)

acs_d = download_acs_vars("WA")
census_d = download_2010_vars("WA")

block_d = download_fcc("data-raw/fcc.csv") %>%
    load_state("WA", .) %>%
    est_2020() %>%
    harmonize_vars(census_d, acs_d)
print(block_d)
```
```{r setup, echo=F, message=F}
library(blockpop)
library(dplyr)
library(stringr)
library(tidyr)

fcc_d = load_state("WA", here::here("data-raw/fcc.csv"))
block_d = est_2020(fcc_d)
acs_d = readr::read_rds(here::here("data-raw/acs.rds"))
census_d = readr::read_rds(here::here("data-raw/census2010.rds"))
block_d = harmonize_vars(block_d, census_d, acs_d)
print(block_d)
```

Then we join summarize our block-level estimates at the block group level and
join them to the ACS data to compare.

```{r}
sum_d = block_d %>%
    select(-starts_with("vap_")) %>%
    mutate(bgroup = str_sub(block, 1, 12)) %>%
    select(-block, -state) %>%
    group_by(bgroup) %>%
    summarize(across(everything(), sum)) %>%
    rename_with(~ str_c("est_", .), .cols=starts_with("pop_")) %>%
    inner_join(rename_with(acs_d, ~ str_c("acs_", .), 
                           .cols=-bgroup), by="bgroup") %>%
    pivot_longer(c(starts_with("est_pop_"), starts_with("acs_pop_")),
                 names_to=c("source", "type", "race"),
                 names_sep="_", values_to="count") %>%
    select(-type) %>%
    pivot_wider(names_from=source, values_from=count) %>%
    mutate(est_pct = est/pop2020,
           acs_pct = acs/acs_pop)
```

Plotting, we see excellent agreement between the ACS and estimated demographic
percentages at the block group level. In fact, the Spearman correlation is 
`r sprintf("%.2f%%", with(sum_d, 100*cor(acs_pct, est_pct, method="spearman", use="complete.obs")))`.

```{r}
library(ggplot2)

ggplot(sum_d, aes(acs_pct, est_pct, color=race)) +
    geom_point(size=0.1, alpha=0.5) +
    coord_cartesian(expand=0) +
    guides(color=guide_legend(override.aes=list(size=5))) +
    theme_bw() 
```

This agreement occurs with 2020 block total population estimates exactly equal
to the projections from `est_2020()`, and using block-level demographic estimates
that are as close as possible to the 2010 Census counts. This means much more
accuracy and granularity than simple population-based disaggregation from
block-group-level ACS data.
